\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,bm,hyperref}
\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{0.8in}
\addtolength{\textheight}{0.8in}
\addtolength{\topmargin}{-.4in}
\newtheoremstyle{quest}{\topsep}{\topsep}{}{}{\bfseries}{}{ }{\thmname{#1}\thmnote{ #3}.}
\theoremstyle{quest}
\newtheorem*{definition}{Definition}
\newtheorem*{theorem}{Theorem}
\newtheorem*{question}{Problem}
\newtheorem*{exercise}{Exercise}
\newtheorem*{challengeproblem}{Challenge Problem}
\newcommand{\name}{%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% put your name here, so we know who to give credit to %%
CS189: Introduction to Machine Learning
}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\hw}{%%%%%%%%%%%%%%%%%%%%
%% and which homework assignment is it? %%%%%%%%%
%% put the correct number below %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
2
}
\newcommand{\duedate}{Due date: }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\vspace{-50pt}
\huge \name
\\\vspace{10pt}
\Large Homework \hw
\\\vspace{10pt}
\large Due: September 24, 2015 at 11:59pm}
\date{}
\author{}

\markright{\name\hfill Homework \hw\hfill}

%% If you want to define a new command, you can do it like this:
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}

%% If you want to use a function like ''sin'' or ''cos'', you can do it like this
%% (we probably won't have much use for this)
% \DeclareMathOperator{\sin}{sin}   %% just an example (it's already defined)


\begin{document}
\maketitle

\begin{picture}(0,0)
\put(-20,180){
\textbf{Name:} \hspace{6cm}
\textbf{Student ID:}  
}
\end{picture}
\vspace{-0.75in}

% =============================================================
% Instructions: 
\section*{Instructions:}
\begin{itemize}
\item Homework 2 is completely a written assignment, no coding involved.
\item Please write (legibly!) or typeset your answers in the space provided. If you choose to typeset your answers, please use this template file (\href{https://bcourses.berkeley.edu/courses/1352792/announcements}{\underline{hw2.tex}}), provided on bCourses announcement page. If there is not enough space for your answer, you can continue your answer on a separate page (for example : You might want to append pages in Questions 6,7,8).
\item Submit a pdf of your answers to \underline{\url{https://gradescope.com}} under Homework 2. A photograph or scanned copy is acceptable as long as it is clear with good contrast. You should be able to see CS 189/289 on gradescope when you login with your primary e-mail address used in bCourses. Please let us know if you have any problems accessing gradescope.
\item While submitting to Gradescope, you will have to select the region containing your answer for each of the question. Thus, write the answer to a question (or given part of the question) at one place only. 
\item Start early and don't wait until last minute to submit the assignment as the submission procedure might take sometime too.
\end{itemize}

\section*{About the Assignment:}
\begin{itemize}
\item This assignment tries to refresh the concepts of probability, linear algebra and matrix calculus. 
\item Questions 1 to 6 are dedicated to deriving fundamental results related to these concepts. You might want to refer your math class textbooks for help.
\item Questions 7,8 discuss few applications of these concepts in machine learning. 
\item Hope you will enjoy doing the assignment !
\end{itemize}
\noindent
\textbf{Homework Party : Sept 21, 2-4pm in the Wozniak Lounge, SODA 430}
\newpage

% =============================================================

\begin{question}[1]
A target is made of 3 concentric circles of radii $1/{\sqrt{3}}$, $1$ and $\sqrt{3}$ feet. Shots within the inner circle are given 4 points, shots within the next ring are given 3 points, and shots within the third ring are given 2 points. Shots outside the target are given 0 points.

Let $X$ be the distance of the hit from the center (in feet), and let the p.d.f of $X$ be
\[
f(x) =
  \begin{cases}
   \frac{2}{\pi (1+x^2)} & x>0 \\
   0 &  \text{otherwise}
  \end{cases}
\]
What is the expected value of the score of a single shot?
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 1 here %%%%%%%%%%%
\newpage

% =============================================================

\begin{question}[2]
Assume that the random variable $X$ has the exponential distribution
\[
f(x|\theta) = \theta e^{-\theta x} \ \ \ \ \ \ \ \ x > 0, \theta > 0
\]
where $\theta$ is the parameter of the distribution. Use the method of maximum likelihood to estimate $\theta$ if 5 observations of $X$ are $x_1 = 0.9$, $x_2 = 1.7$, $x_3 = 0.4$, $x_4 = 0.3$, and $x_5 = 2.4$, generated i.i.d. (i.e., independent and identically distributed).
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 2 here %%%%%%%%%%%
\newpage

% =============================================================

\begin{question}[3]
The polynomial kernel is defined to be
\[
k(\bf{x},\bf{y}) = (\bf{x}^T\bf{y} + c)^d
\]
where $\bf{x},\bf{y} \in \mathbb{R}^n$, and $c \geq 0$. When we take $d=2$, this kernel is called the quadratic kernel.
\begin{itemize}
\item[(a)] Find the feature mapping $\Phi(\bf{z})$ that corresponds to the quadratic kernel.
\item[(b)] How do we find the optimal value of $d$ for a given dataset?
\end{itemize}
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 3 (a) here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 3 (b) here %%%%%%%%%%%
\newpage

% =============================================================

\noindent
\textbf{Def}: Let $A \in \mathbb{R}^{n \times n}$ be a symmetric matrix. We say that $A$ is positive definite if $\forall x\in \mathbb{R}^n,\ x^\top Ax > 0$. Similarly, we say that $A$ is positive semidefinite if $\forall x \in \mathbb{R}^n,\ x^\top Ax \geq 0$.
\begin{question}[4]
Let $x = \begin{bmatrix}x_1 & \cdots & x_n\end{bmatrix}^\top \in \mathbb{R}^n$, and let $A \in \mathbb{R}^{n \times n}$ be the square matrix
\begin{equation*}
A = \begin{bmatrix}
  a_{11} & a_{12} & \cdots & a_{1n} \\
  a_{21} & a_{22} & \cdots & a_{2n} \\
  \vdots  & \vdots  & \ddots & \vdots  \\
  a_{n1} & a_{n2} & \cdots & a_{nn}
\end{bmatrix}
\end{equation*}

\begin{itemize}
\item[(a)] Give an explicit formula for $x^\top A x$. Write your answer as a sum involving the elements of $A$ and $x$. 
\item[(b)] Show that if $A$ is positive definite, then the entries on the diagonal of $A$ are positive (that is, $a_{ii} > 0$ for all $1 \leq i \leq n$).
\end{itemize}
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 4a here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 4b here %%%%%%%%%%%


% =============================================================

\newpage
\begin{question}[5]
Let $B$ be a positive semidefinite matrix. Show that $B + \gamma I$ is positive definite for any $\gamma > 0$.
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 5 here %%%%%%%%%%%
\newpage

% =============================================================

\begin{question}[6 : Derivatives and Norms]
Derive the expression for following questions. Do not write the answers directly.
\begin{itemize}
\item[(a)] Let $\textbf{x}, \textbf{a} \in \R^n$. Derive $\frac{\partial \left(\textbf{x}^T\textbf{a}\right)}{\partial \textbf{x}}$.
\item[(b)] Let \textbf{A} be a $n\times n$ matrix and $\textbf{x}$ be a vector in $\R^n$. Derive $\frac{\partial \left(\textbf{x}^T\textbf{A}\textbf{x} \right)}{\partial \textbf{x}}$.
\item[(c)] Let \textbf{A}, \textbf{X} be $n\times n$ matrices. Derive $\frac{\partial \text{Trace(\textbf{XA})}}{\partial \textbf{X}}$.
\item[(d)] Let \textbf{X} be a $m\times n$ matrix, $\textbf{a}\in \R^m$ and $\textbf{b}\in \R^n$. Derive $\frac{\partial \left(\textbf{a}^T\textbf{Xb}\right)}{\partial \textbf{X}}$.
\item[(e)] Let $\textbf{x}\in \R^n$. Prove that $\|\textbf{x}\|_2 \leq \|\textbf{x}\|_1 \leq \sqrt{n}\|\textbf{x}\|_2$. Here $\|x\|_2=\sqrt{\sum_{i=1}^{n} x_i^2}$ and $\|\textbf{x}\|_1=\sum_{i=1}^{n} |x_i|$.
\end{itemize}
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 6a here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 6b here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 6c here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 6d here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 6e here %%%%%%%%%%%

% =============================================================

\newpage
\begin{question}[7 : Application of Matrix Derivatives]
\mbox{}\\
\noindent
Let \textbf{X} be a $n\times d$ data matrix, $\textbf{Y}$ be the corresponding $n\times 1$ target/label matrix and $\boldsymbol{\Lambda}$ be the diagonal $n\times n$ matrix containing weight of each example. 
Expanding them, we have $\textbf{X} = \left[ \begin{smallmatrix} (\textbf{x}^{(1)})^T \\ (\textbf{x}^{(2)})^T \\ \dots \\ (\textbf{x}^{(n)})^T \end{smallmatrix} \right]$, 
$\textbf{Y} = \left[ \begin{smallmatrix} \textbf{y}^{(1)} \\ \textbf{y}^{(2)} \\ \dots \\ \textbf{y}^{(n)} \end{smallmatrix} \right]$ 
and $\boldsymbol{\Lambda} = \text{diag}(\lambda^{(1)},\lambda^{(2)},\dots,\lambda^{(n)})$\\
where $\textbf{x}^{(i)} \in \mathbb{R}^d$, $\textbf{y}^{(i)} \in \mathbb{R}$, and $\lambda^{(i)} > 0$ 
$\;\;\forall\;\;i\in\{1\dots n\}$. \textbf{X}, \textbf{Y} and $\boldsymbol{\Lambda}$ are fixed and known.\\
In the remaining parts of this question, we will try to fit a weighted linear regression model for this data. We want to find the value of weight vector $w$ which best satisfies the following equation $y^{(i)}=\textbf{w}^T\textbf{x}^{(i)}+\epsilon^{(i)}$ where $\epsilon$ is noise. This is achieved by minimizing the weighted noise for all the examples. Thus, our risk function is defined as follows:
\begin{align*}
R[\textbf{w}] &= \sum_{i=1}^{n} \lambda^{(i)}(\epsilon^{(i)})^2\\
&= \sum_{i=1}^{n} \lambda^{(i)}(\textbf{w}^T\textbf{x}^{(i)}-y^{(i)})^2
\end{align*}
\begin{itemize}
\item[(a)] Write this risk function $R[\textbf{w}]$ in matrix notation, i.e., in terms of \textbf{X}, \textbf{Y}, $\boldsymbol{\Lambda}$ and \textbf{w}.
\item[(b)] Find the value of \textbf{w}, in matrix notation, that minimizes the risk function obtained in Part (a). You can assume that $\textbf{X}^T\boldsymbol{\Lambda}\textbf{X}$ is full rank matrix. Hint: You can use the expression derived in Q-6(b).
\item[(c)] What will be the answer for questions in Parts (a) and (b) if you add $\text{L}_2$ regularization (i.e., shrinkage) on \textbf{w}? The L2 regularized risk function, for $\gamma>0$, is
\begin{align*}
R[\textbf{w}] &= \sum_{i=1}^{n} \lambda^{(i)}(\textbf{w}^T\textbf{x}^{(i)}-y^{(i)})^2 + \gamma \|\textbf{w}\|^2_2
\end{align*}
Hint: You can make use of the result in Q-5.
\item[(d)] What role does the regularization (i.e., shrinkage) play in fitting the regression model and how ? You can observe the difference in expressions for $\textbf{w}$ obtained in Parts (c) and (d), and argue.
\end{itemize}
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 7a here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 7b here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 7c here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 7d here %%%%%%%%%%%
\newpage

% =============================================================

\begin{question}[8: Classification]
Suppose we have a classification problem with classes labeled $1, \dotsc, c$ and an additional doubt category labeled as $c+1$.
Let the loss function be the following:\\
\[
\ell(f(x) = i, y = j) =
  \begin{cases}
   0 &  \mathrm{if}\ i=j \quad i,j\in\{1,\dotsc,c\} \\
   \lambda_r       & \mathrm{if}\ i=c+1 \\
   \lambda_s       & \text{otherwise}
  \end{cases}
\]
where $\lambda_r$ is the loss incurred for choosing doubt and $\lambda_s$ is the loss incurred for making a misclassification. Note that $\lambda_r \ge 0$ and $\lambda_s \ge 0$. \\
Hint : The risk of classifying a new datapoint as class $i\in\{1,2,\dots,c+1\}$ is $$R(\alpha_i|x) = \sum_{j=1}^{j=c} \ell(f(x) = i, y = j) P(\omega_j|x)$$

\begin{itemize}
\item[(a)] Show that the minimum risk is obtained if we follow this policy: (1) choose class~$i$ if $P(\omega_i|x) \geq P(\omega_j|x)$ for all $j$ and $P(\omega_i|x) \geq 1-\lambda_r/\lambda_s$, and (2) choose doubt otherwise.
\item[(b)] What happens if $\lambda_r=0$? What happens if $\lambda_r>\lambda_s$?
\end{itemize}
\end{question}
\textbf{Solution:}
%%%%%%%%%%% Write your solution to problem 8a here %%%%%%%%%%%
%%%%%%%%%%% Write your solution to problem 8b here %%%%%%%%%%%
\newpage

% =============================================================

\end{document}
